Bot's Guide to Being Human: Investigating the Effect of Conversation Architecture Elements on the Anthropomorphized Perceptions of Conversational Agents

### 2023-02-12 paper.tex
first draft up to findings; updated perception table

\usepackage[linewidth=1pt]{mdframed}
\usepackage{lipsum}
\usepackage{multirow}
\usepackage{graphicx}
\usepackage{booktabs}

@@

\section{Introduction}

@@

\subsection{Search Procedure}

@@

\begin{mdframed}
"conversational agent" OR "natural language interface" OR "IPA" OR "intelligent personal assistant" OR "chatbot" OR "speech interface" OR "voice assistant" OR "intelligent agent" OR "human-chatbot communication" OR "virtual agent" OR "dialog* system" OR "voice user interface" OR "human computer dialog*"
\end{mdframed}

\subsection{Selection Criteria}
The initial query retrieved 2901 unique publications. The following selection criteria are applied to identify literature related to the effect of conversational architecture on perception of agents:
\begin{itemize}
  \item Peer-reviewed publications written in English
  \item Voice-based or text-based conversational agents
  \item Experiment study on the impact of conversation architecture elements on the anthropomorphized perceptions of agents
  \item Impact of conversation architecture elements can be separated from other effects (e.g. embodiment)

@@

\begin{figure}[h]
  \centering
  \includegraphics[width=0.6\columnwidth]{fig-prisma.png}
  \caption{Literature search using PRISMA guidelines}
  \label{fig:prisma}
\end{figure}

@@

The papers reviewed (n=57) were published between May 2011 and November 2022. More than 75\% of the papers were published in or after 2019, with a slight trend upwards over the years (see Figure \ref{fig:paper}). The vast majority of the papers in our corpus were published in conference proceedings (n=50), with the remainder papers published in journals. Out of the papers published in conferences, the top conferences were The ACM Conference on Human Factors in Computing Systems (n=16), Conversational User Interfaces (n=7), Human-Agent Interaction (n=4), and Intelligent Virtual Agents (n=4). Out of the papers published in journals, some venues include the International Journal of Human-Computer Interaction (n=2) and Interacting with Computers (n=1).

@@

\begin{figure*}[h]
  \centering
  \includegraphics[width=\textwidth]{fig-paper.png}
  \caption{Stacked barchart of number of papers by year and venue}
  \label{fig:paper}
\end{figure*}

@@

\subsubsection*{Anthropomorphized Perceptions}
The exact quotes of anthropomorphized perception measurements were extracted from each paper. Inductive thematic analysis was used to group these verbatim quotes into codes based on their meanings. For example, the code likeability of the agent contains the questionnaire item "this voice agent was likeable" used by \cite{cuadra2021my}\cmt{[67]}, Godspeed questionnaire \cite{bartneck2009measurement}\cmt{godspeed}'s set of questions on likeability used by \cite{linnemann2018can}\cmt{[15]}, and Subjective Assessment of Speech System Interface (SASSI) questionnaire \cite{hone2000towards}\cmt{sassi}'s set of questions on likeability used by  \cite{chan2021kinvoices}\cmt{[74]}\cite{choi2020nobody}\cmt{[54]}. This step resulted in 83 unique codes. The codes are then organized based similarity of the meanings, resulting in 11 basic themes. For example, the basic theme of \textit{personality traits} of the agent contains measurements such as likeability, friendliness, and warmth. Lastly, the 11 basic themes are categorized into 4 organizing themes: \textit{perception of interaction with agent, perception of agent's ability, perception of social connection with agent, and perception of agent's humanness} (Table \ref{tab:perceptions}).

\subsubsection*{Conversation Architecture Elements}
The conversation architecture elements used in the study were extracted from each paper. Some of the papers were applying multiple elements to the CA design, such as the the anthropomorphic design used in this study \cite{seeger2021chatbots}\cmt{[35]}. In the case where multiple elements were used, the codes broke down the conversation architecture into its core components. For example, the anthropomorphic design used in Seeger et al's study \cite{seeger2021chatbots}\cmt{[35]} was broken down to the elements of emotional expressions, is-typing indicator, emoticons, and response delay to capture as codes. 58 unique codes were created as part of this process, capturing elements like sentiment-adaptive responses \cite{diederich2019emulating}\cmt{[25]}, lexical alignment \cite{spillner2021talk}\cmt{[18]}, and typos \cite{westerman2019believe}\cmt{[9]}. Codes were then organized based on similarity of the elements. For example, the basic theme of \textit{disfluency} contains elements of fillers \cite{jeong2019exploring}\cmt{[10]}\cite{wester2015artificial}\cmt{[14]}, interjections \cite{ceha2022expressive}\cmt{[77]}\cite{hu2021enhancing}\cmt{[56]}, and repetitions \cite{yang2021effect}\cmt{[72]}. Lastly, these basic themes are then categorized into organizing themes adapted from the social cues taxonomy structure by Feine et al \cite{feine2019taxonomy}. Based on this paper, the Verbal-Content category is mapped to our organizing theme of \textit{linguistic content}, and the Verbal-Style category is mapped to our organizing theme of \textit{linguistic style}.
Given our analysis focuses on the text and voice aspects of a conversational agent and does not analyze the embodiment aspects of design, the visual category is not applicable for our analysis. Also, we collapsed the auditory and invisible categories into the organizing theme of \textit{non-linguistic format} (Table \ref{tab:cues}).

\subsubsection*{Relationship Between Anthropomorphized Perceptions and Conversation Architecture Elements}
Based on the codes developed for anthropomorphized perceptions and conversation architecture elements above, we extracted the relationships between perceptions and architecture elements from each paper, noting whether the architecture element had an impact of on the perceptions. Out of the 57 papers, 264 connections between anthropomorphized perceptions and conversation architecture elements were found. To analyze specifically the relationships between the specifics of conversation architecture that affect the anthropomorphized perceptions of CAs, 69 connections that did not result in significant relationships were discarded, resulting in 195 relationships for analysis. For example, the specific connection between style matching and user satisfaction was discarded because Hoegen et al. \cite{hoegen2019end}\cmt{[31]} did not find a significant difference between the style matching agent and the non-style matching agent for overall interaction satisfaction. Out of the 195 relationships, 185 were relationships between individual architecture elements and perceptions. These single architecture element to perception relationships are visualized as a heatmap based on literature coverage within the literature review corpus, as seen in Figure \ref{fig:heatmap-impact}. Perceptions impacted by the composite of multiple architecture elements in an agent will be discussed in more details in the Composite Architecture Elements section.

@@

\begin{table*}[]
\resizebox{\textwidth}{!}{%
\begin{tabular}{@{}lllc@{}}
\toprule
\textbf{Category} & \textbf{Theme}   & \textbf{Sample Codes}                                  & \textbf{No. Papers} \\ \midrule
\multirow{3}{*}{Perception of Interaction with Agent}    & Usability       & accuracy, ease of use, efficiency, helpfulness & 23 \\
 & Engagement         & enjoyment, annoyance, desirable, intention to use  & 28         \\
 & Satisfaction       & service satisfaction, quality of interaction       & 12         \\ \midrule
\multirow{3}{*}{Perception of Agent's Ability}              & Intelligence    & knowledgeable, intelligent, expertise                             & 11 \\
 & Competence         & competent, capable                                 & 4          \\
 & Trust              & credibility, trustworthy, truthfulness, confidence & 15         \\ \midrule
\multirow{3}{*}{Perception of Social Connection with Agent}  & Conversation Tone  & appropriate, expressive, empathetic, persuasive    & 13 \\
 & Social Presence & connectedness, familiarity, similarity, psychological distance    & 18        \\
 & Intimacy           & intimate, rapport, quality of relationship         & 7          \\ \midrule
\multirow{2}{*}{Perception of Agent's Humanness}            & Human-likeness  & human-like, natural, artificial, machine-like                     & 20 \\
 & Personality Traits & friendly, kind, warm, creepy, likeable, polite     & 27         \\ \bottomrule
\end{tabular}%
}
\caption{Perception measures}
\label{tab:perceptions}
\end{table*}

@@

\textbf{Perception of interaction with a CA} evaluates the overall interaction quality users had with the conversational agent. The three basic themes for this perceptions are usability, engagement and satisfaction. \textit{Usability} measures the utilitarian aspects of the interaction, whether it was accurate, easy to use, efficient or helpful. Some commonly used methods to measure usability include the response accuracy portion of the SASSI questionnaire \cite{hone2000towards}\cmt{sassi}, or the NASA Task Load Index (NASA-TLX) \cite{hart1988development}\cmt{nasa} to measure cognitive workload. Questions such as "the system is easy to use" or "it is easy to understand the agent" are also measures of usability. \textit{Engagement} on the other hand measures users' emotional reactions, whether they have enjoyed the conversation with CA, or annoyed or frustrated with the interaction. Some commonly used methods to measure engagement include  the annoyance portion of the SASSI questionnaire \cite{hone2000towards}\cmt{sassi}, and the Use Engagement Scale (UES) \cite{o2018practical}\cmt{ues}. Questions such as "I enjoyed using the system" or "I felt frustrated with the agent" are also measures of engagement. Lastly, \textit{Satisfaction} measures users' overall satisfaction interacting with the agent. Questions such as "the overall assessment of conversing with the CA was satisfactory" are used to measure this perception.

\textbf{Perception of a CA's ability} evaluates the perceived capabilities of the agent. While there are some elements of system performance in this measure, the main aspect we are interested in is the perception differences for agents with identify system capabilities but with different architecture elements. Specifically, \textit{intelligence} measures the agent's expertise and knowledge, and it is commonly administered through the Godspeed questionnaire \cite{bartneck2009measurement}\cmt{godspeed} on the set of questions related to perceived intelligence. Also survey questions such as asking users about the agent's intelligence and domain knowledge are used to measure perceived intelligence. \textit{Competence} takes intelligence one step further by evaluating the agent's ability to put its intelligence into practice. To measure a CA's competence, there are usually surveys or qualitative feedback that the agent is capable and competent, and that the users have confidence in the agent's ability to get the job done. Lastly, \textit{trust} evaluates agent's truthfulness, credibility and benevolence. Some commonly used methods to evaluate trust include the Trust Propensity Scale \cite{mayer1999effect} and the Individualized Trust Scale (ITS) \cite{wheeless1977measurement}. Questions such as "is the agent honest" and "can I trust the agent with sensitive information" are also measures of trust.

\textbf{Perception of the social connection with a CA} evaluates the emotional connections that users have with agents. The three basic themes of conversation tone, social presence and intimacy are included in this category. \textit{Conversation tone} measure the affective impression of the agent's tone, such as empathy, expressiveness, or passionate. Also, it measures whether the tone used by the agent is persuasive or appropriate for the conversation. \textit{Social presence} measures the sense of connectness and psychological distance users has with an agent. Some aspects in this measure include a sense of familiarity or similarity with the agent, whether users feel the agent behaves like them or have similar attitudes to them. The measure of \textit{Intimacy} extends social presence into the realm of the quality of relationships with an agent. Some commonly used measures include the set of social attraction questions from the interpersonal attraction questionnaire \cite{mccroskey1975development} and the quality of relationship inventory (QRI) \cite{pierce1997assessing}. These questionnaires include questions like "I think the agent could be a friend of mine", or "I feel we could establish a personal relationship with each other".

\textbf{Perception of a CA's humanness} evaluates human-likeness and the anthropomorphized personality traits that users assign to agents. \textit{Human-likeness} measures whether the agent presented itself as natural and human-like, or artificial and machine-like. The Godspeed questionnaire \cite{bartneck2009measurement}\cmt{godspeed} set of questions related to anthropomorphism and the Ascent of Man scale \cite{kteily2015ascent} are commonly used methods to assess human-likeness. Survey questions with semantic scales such as "human-like / machine-like" and "artificial / natural" are also used for this perception. \textit{Personality traits} captures the human characteristics that are attributes to the agent, such as warm, friendly, likeable, and polite. This is commonly captured in the qualitative feedback from users, commenting on whether the agent is extroverted or introverted, or the human characteristics they perceive the agent such as funny or witty. Some studies used the measure of the Big-5 personality traits \cite{gosling2003very} to map an agent's disposition on various personality dimensions.

@@

\section{Framework}
%Wrote based on papers up to [60], 25 papers

While the number of papers studying voice (n=27) vs. text (n=30) modalities is not significantly different, there are some interesting trends in the perceptions studied for CAs with different modalities. Overall, there are less perceptions of agent studied for text-based agents (n=101) as compared to voice-based agents (n=163).

This is especially evident in the perception of social connection with agent and perception of agent's humanness. For the perception of social connection with a CA, there are only 19 perceptions measured for text-based agents, while there are 35 perceptions measured for voice-based agents. Similarly for perception of agent's humanness, there are 36 perceptions measured for text-based agents, as compared to 52 perceptions of voice-based agents. This highlights that there is a

>>>>

\usepackage[linewidth=1pt]{mdframed}
\usepackage{lipsum}
\usepackage{multirow}
\usepackage{graphicx}
\usepackage{booktabs}
\usepackage{makecell}

@@

\section{Introduction (Not Ready)}

@@

\subsection{Data Collection}

@@

\textit{"conversational agent" OR "natural language interface" OR "IPA" OR "intelligent personal assistant" OR "chatbot" OR "speech interface" OR "voice assistant" OR "intelligent agent" OR "human-chatbot communication" OR "virtual agent" OR "dialog* system" OR "voice user interface" OR "human computer dialog*"}
\newline

The initial query retrieved 2901 unique publications. The following selection criteria are applied to identify literature related to the effect of conversational architecture on perception of agents:
\begin{itemize}
  \item Peer-reviewed publications written in English
  \item Voice-based or text-based conversational agents
  \item Experiment study on the impact of conversation architecture elements on the anthropomorphized perceptions of agents
  \item Impact of conversation architecture elements can be separated from other effects (e.g. embodiment)

@@

\begin{figure}[h]
  \centering
  \includegraphics[width=1\columnwidth]{fig-prisma.png}
  \caption{Literature search using PRISMA guidelines}
  \label{fig:prisma}
\end{figure}

@@

The papers reviewed (n=57) were published between May 2011 and November 2022. More than 75\% of the papers were published in or after 2019, with a slight trend upwards over the years (see Figure \ref{fig:paper}). The vast majority of the papers in our corpus were published in conference proceedings (n=50), with the remainder papers published in journals. Out of the papers published in conferences, the top conferences were The ACM Conference on Human Factors in Computing Systems\footnote{https://dl.acm.org/conference/chi} (n=16), Conversational User Interfaces\footnote{https://dl.acm.org/conference/cui} (n=7), Human-Agent Interaction\footnote{https://dl.acm.org/conference/hai} (n=4), and Intelligent Virtual Agents\footnote{https://dl.acm.org/conference/iva} (n=4). 
%Out of the papers published in journals, some venues include the International Journal of Human-Computer Interaction\footnote{https://www.tandfonline.com/journals/hihc20} (n=2) and Interacting with Computers\footnote{https://academic.oup.com/iwc} (n=1).

@@

\begin{figure}[h]
  \centering
  \includegraphics[width=\columnwidth]{fig-paper.png}
  \caption{Number of papers by year}
  \label{fig:paper}
\end{figure}

@@

\subsubsection*{Anthropomorphized Perceptions} To study RQ1, the exact quotes of anthropomorphized perception measurements were extracted from each paper. Inductive thematic analysis was used to group these verbatim quotes into codes based on their meanings. For example, the code likeability of the agent contains the questionnaire item "this voice agent was likeable" used by \cite{cuadra2021my}\cmt{[67]}, Godspeed questionnaire \cite{bartneck2009measurement}\cmt{godspeed}'s set of questions on likeability used by \cite{linnemann2018can}\cmt{[15]}, and Subjective Assessment of Speech System Interface (SASSI) questionnaire \cite{hone2000towards}\cmt{sassi}'s set of questions on likeability used by  \cite{chan2021kinvoices}\cmt{[74]}\cite{choi2020nobody}\cmt{[54]}. This step resulted in 83 unique codes. The codes are then organized based similarity of the meanings, resulting in 11 basic themes. For example, the basic theme of \textit{personality traits} of the agent contains measurements such as likeability, friendliness, and warmth. Lastly, the 11 basic themes are categorized into 4 organizing themes: \textit{perception of interaction with agent, perception of agent's ability, perception of social connection with agent, and perception of agent's humanness} (Table \ref{tab:perceptions}).

\subsubsection*{Conversation Architecture Elements}
The study RQ2, the conversation architecture elements used in the study were extracted from each paper. Some of the papers were applying multiple elements to the CA design, such as the the anthropomorphic design used in this study \cite{seeger2021chatbots}\cmt{[35]}. In the case where multiple elements were used, the codes broke down the conversation architecture into its core components. For example, the anthropomorphic design used in Seeger et al's study \cite{seeger2021chatbots}\cmt{[35]} was broken down to the elements of emotional expressions, is-typing indicator, emoticons, and response delay to capture as codes. 58 unique codes were created as part of this process, capturing elements like sentiment-adaptive responses \cite{diederich2019emulating}\cmt{[25]}, lexical alignment \cite{spillner2021talk}\cmt{[18]}, and typos \cite{westerman2019believe}\cmt{[9]}. Codes were then organized based on similarity of the elements. For example, the basic theme of \textit{disfluency} contains elements of fillers \cite{jeong2019exploring}\cmt{[10]}\cite{wester2015artificial}\cmt{[14]}, interjections \cite{ceha2022expressive}\cmt{[77]}\cite{hu2021enhancing}\cmt{[56]}, and repetitions \cite{yang2021effect}\cmt{[72]}. Lastly, these basic themes are then categorized into organizing themes adapted from the social cues taxonomy structure by Feine et al \cite{feine2019taxonomy}. Based on this paper, the Verbal-Content category is mapped to our organizing theme of \textit{linguistic content}, and the Verbal-Style category is mapped to our organizing theme of \textit{linguistic style}.
Given our analysis focuses on the text and voice aspects of a conversational agent and does not analyze the embodiment aspects of design, the visual category is not applicable for our analysis. Also, we collapsed the auditory and invisible categories into the organizing theme of \textit{non-linguistic format} (Table \ref{tab:cues}).

\subsubsection*{Relationship Between Anthropomorphized Perceptions and Conversation Architecture Elements}
For RQ3, we extracted the relationships between perceptions and architecture elements from each paper based on the codes developed for anthropomorphized perceptions and conversation architecture elements above, noting whether the architecture element had an impact of on the perceptions. Out of the 57 papers, 264 connections between anthropomorphized perceptions and conversation architecture elements were found. To analyze specifically the relationships between the specifics of conversation architecture that affect the anthropomorphized perceptions of CAs, 69 connections that did not result in significant relationships were discarded, resulting in 195 relationships for analysis. For example, the specific connection between style matching and user satisfaction was discarded because Hoegen et al. \cite{hoegen2019end}\cmt{[31]} did not find a significant difference between the style matching agent and the non-style matching agent for overall interaction satisfaction. Out of the 195 relationships, 185 were relationships between individual architecture elements and perceptions. These single architecture element to perception relationships are visualized as a heatmap based on literature coverage within the literature review corpus, as seen in Figure \ref{fig:heatmap-impact}. Perceptions impacted by the composite of multiple architecture elements in an agent will be discussed in more details in the Composite Architecture Elements section.

@@

\begin{table*}[ht]
\renewcommand*{\arraystretch}{1.4}
\resizebox{\textwidth}{!}{%
\begin{tabular}{@{}p{3.1cm} | p{2cm} | p{4cm} | p{1.5cm} | p{6cm} @{}}
%!{\vrule width 1.2pt} 
\Xhline{1.2pt}
\textbf{Category} & \textbf{Theme} & \textbf{Sample Codes} & \textbf{No. Papers} & \textbf{Papers} 
\\ \Xhline{1.2pt}
\multirow{3}{*}{\parbox{3cm}{Perception of Interaction with Agent}}
    & Usability & accuracy, ease of use, efficiency, helpfulness & 23 
    & \cite{ashktorab2019resilient}\cmt{[88]}
    \cite{ceha2022expressive}\cmt{[77]}
    \cite{chan2021kinvoices}\cmt{[74]}
    \cite{choi2020nobody}\cmt{[54]}
    \cite{cuadra2021my}\cmt{[67]}
    \cite{dubiel2020persuasive}\cmt{[60]}
    \cite{elsholz2019exploring}\cmt{[61]}
    \cite{haas2022keep}\cmt{[78]}
    \cite{habler2019effects}\cmt{[63]}
    \cite{healey2013relating}\cmt{[39]}
    \cite{hu2022polite}\cmt{[76]}
    \cite{huiyang2022improving}\cmt{[17]}
    \cite{jestin2022effects}\cmt{[81]}
    \cite{kim2019comparing}\cmt{[89]}
    \cite{kim2020can}\cmt{[24]}
    \cite{kraus2020effects}\cmt{[64]}
    \cite{linnemann2018can}\cmt{[15]}
    \cite{ma2022ask}\cmt{[29]}
    \cite{miehle2018exploring}\cmt{[51]}
    \cite{misu2011toward}\cmt{[83]}
    \cite{roy2021users}\cmt{[71]}
    \cite{spillner2021talk}\cmt{[18]}
    \cite{wilhelm2022keep}\cmt{[28]}
\\ \cline{2-5}
    & Engagement & enjoyment, annoyance, desirable, intention to use  & 28
    & \cite{andrews2012system}\cmt{[38]}
    \cite{ceha2022expressive}\cmt{[77]}
    \cite{ceha2021can}\cmt{[57]}
    \cite{chan2021kinvoices}\cmt{[74]}
    \cite{cox2022does}\cmt{[27]}
    \cite{cuadra2021my}\cmt{[67]}
    \cite{elsholz2019exploring}\cmt{[61]}
    \cite{fadhil2018effect}\cmt{[52]}
    \cite{gnewuch2022opposing}\cmt{[20]}
    \cite{go2021conversational}\cmt{[80]}
    \cite{healey2013relating}\cmt{[39]}
    \cite{huiyang2022improving}\cmt{[17]}
    \cite{jestin2022effects}\cmt{[81]}
    \cite{kim2020can}\cmt{[24]}
    \cite{kim2019comparing}\cmt{[89]}
    \cite{lee2020hear}\cmt{[23]}
    \cite{linnemann2018can}\cmt{[15]}
    \cite{ma2022ask}\cmt{[29]}
    \cite{miehle2018exploring}\cmt{[51]}
    \cite{miyamoto2017improving}\cmt{[46]}
    \cite{moilanen2022measuring}\cmt{[82]}
    \cite{roy2021users}\cmt{[71]}
    \cite{spillner2021talk}\cmt{[18]}
    \cite{volkel2021examining}\cmt{[69]}
    \cite{volkel2022user}\cmt{[75]}
    \cite{xiao2021let}\cmt{[73]}
    \cite{yang2021effect}\cmt{[72]}
    \cite{zhu2022effects}\cmt{[26]}    
\\ \cline{2-5}
    & Satisfaction & service satisfaction, quality of interaction & 12
    & \cite{ceha2022expressive}\cmt{[77]}
    \cite{choi2020nobody}\cmt{[54]}
    \cite{diederich2019emulating}\cmt{[25]}
    \cite{elsholz2019exploring}\cmt{[61]}
    \cite{habler2019effects}\cmt{[63]}
    \cite{gnewuch2018faster}\cmt{[19]}
    \cite{hoegen2019end}\cmt{[31]}
    \cite{hu2022polite}\cmt{[76]}
    \cite{ma2022ask}\cmt{[29]}
    \cite{roy2021users}\cmt{[71]}
    \cite{wilhelm2022keep}\cmt{[28]}
    \cite{yang2017perceived}\cmt{[44]}
\\ \Xhline{1.2pt} 
\multirow{3}{*}{\parbox{3cm}{Perception of Agent's Ability}}
    & Intelligence & knowledgeable, intelligent, expertise & 11
    & \cite{ashktorab2019resilient}\cmt{[88]}
    \cite{ceha2021can}\cmt{[57]}
    \cite{chan2021kinvoices}\cmt{[74]}
    \cite{cuadra2021my}\cmt{[67]}
    \cite{dubiel2020persuasive}\cmt{[60]}
    \cite{feijoo2021effects}\cmt{[70]}
    \cite{hu2021enhancing}\cmt{[56]}
    \cite{jeong2019exploring}\cmt{[10]}
    \cite{spillner2021talk}\cmt{[18]}
    \cite{volkel2022user}\cmt{[75]}
    \cite{yang2017perceived}\cmt{[44]} 
\\ \cline{2-5}
    & Competence & competent, capable & 4 
    & \cite{cox2022does}\cmt{[27]}
    \cite{kraus2020effects}\cmt{[64]}
    \cite{lee2019s}\cmt{[55]}
    \cite{westerman2019believe}\cmt{[9]}
\\ \cline{2-5}
    & Trust & credibility, trustworthy, truthfulness, confidence & 15
    & \cite{andrews2012system}\cmt{[38]}
    \cite{chan2021kinvoices}\cmt{[74]}
    \cite{dubiel2020persuasive}\cmt{[60]}
    \cite{fadhil2018effect}\cmt{[52]}
    \cite{healey2013relating}\cmt{[39]}
    \cite{hoegen2019end}\cmt{[31]}
    \cite{huiyang2022improving}\cmt{[17]}
    \cite{jestin2022effects}\cmt{[81]}
    \cite{kraus2020effects}\cmt{[64]}
    \cite{lee2020hear}\cmt{[23]}
    \cite{linnemann2018can}\cmt{[15]}
    \cite{ma2022ask}\cmt{[29]}
    \cite{seeger2021chatbots}\cmt{[35]}
    \cite{tolmeijer2021female}\cmt{[62]}
    \cite{wilhelm2022keep}\cmt{[28]}    
\\ \Xhline{1.2pt} 
\multirow{3}{*}{\parbox{3cm}{Perception of Social Connection with Agent}}
    & Conversation Tone & appropriate, expressive, empathetic, persuasive & 13 
    & \cite{chan2021kinvoices}\cmt{[74]}
    \cite{cox2022does}\cmt{[27]}
    \cite{cuadra2021my}\cmt{[67]}
    \cite{daher2020empathic}\cmt{[58]}
    \cite{diederich2019emulating}\cmt{[25]}
    \cite{dubiel2020persuasive}\cmt{[60]}
    \cite{hu2021enhancing}\cmt{[56]}
    \cite{jestin2022effects}\cmt{[81]}
    \cite{misu2011toward}\cmt{[83]}
    \cite{tolmeijer2021female}\cmt{[62]}
    \cite{yang2021effect}\cmt{[72]}
    \cite{yang2017perceived}\cmt{[44]}
    \cite{zhu2022effects}\cmt{[26]}           
\\ \cline{2-5}
    & Social Presence & connectedness, familiarity, similarity, psychological distance & 18
    & \cite{ceha2022expressive}\cmt{[77]}
    \cite{ceha2021can}\cmt{[57]}
    \cite{chan2021kinvoices}\cmt{[74]}
    \cite{choi2020nobody}\cmt{[54]}
    \cite{cuadra2021my}\cmt{[67]}
    \cite{diederich2019emulating}\cmt{[25]}
    \cite{gnewuch2018faster}\cmt{[19]}
    \cite{gnewuch2018chatbot}\cmt{[21]}
    \cite{gnewuch2022opposing}\cmt{[20]}
    \cite{go2021conversational}\cmt{[80]}
    \cite{khooshabeh2011does}\cmt{[37]}
    \cite{kim2020can}\cmt{[24]}
    \cite{lee2019s}\cmt{[55]}
    \cite{lubis2019positive}\cmt{[43]}
    \cite{lubold2016effects}\cmt{[86]}
    \cite{ma2022ask}\cmt{[29]}
    \cite{niewiadomski2013laugh}\cmt{[85]}
    \cite{westerman2019believe}\cmt{[9]}    
\\ \cline{2-5}
    & Intimacy & intimate, rapport, quality of relationship & 7 
    & \cite{choi2020nobody}\cmt{[54]}
    \cite{khooshabeh2011does}\cmt{[37]}
    \cite{kim2020can}\cmt{[24]}
    \cite{lee2020hear}\cmt{[23]}
    \cite{linnemann2018can}\cmt{[15]}
    \cite{lubold2016effects}\cmt{[86]}
    \cite{westerman2019believe}\cmt{[9]}
\\ \Xhline{1.2pt}
\multirow{2}{*}{\parbox{3cm}{Perception of Agent's Humanness}}
    & Human-likeness & human-like, natural, artificial, machine-like & 20
    & \cite{ashktorab2019resilient}\cmt{[88]}
    \cite{ceha2021can}\cmt{[57]}
    \cite{chan2021kinvoices}\cmt{[74]}
    \cite{choi2020nobody}\cmt{[54]}
    \cite{cox2022does}\cmt{[27]}
    \cite{diederich2019emulating}\cmt{[25]}
    \cite{gnewuch2018faster}\cmt{[19]}
    \cite{haas2022keep}\cmt{[78]}
    \cite{hu2022polite}\cmt{[76]}
    \cite{jeong2019exploring}\cmt{[10]}
    \cite{jestin2022effects}\cmt{[81]}
    \cite{lubis2019positive}\cmt{[43]}
    \cite{ma2022ask}\cmt{[29]}
    \cite{misu2011toward}\cmt{[83]}
    \cite{niewiadomski2013laugh}\cmt{[85]}
    \cite{ouchi2019should}\cmt{[59]}
    \cite{seeger2021chatbots}\cmt{[35]}
    \cite{wester2015artificial}\cmt{[14]}
    \cite{westerman2019believe}\cmt{[9]}
    \cite{zhu2022effects}\cmt{[26]}    
\\ \cline{2-5}
    & Personality Traits & friendly, kind, warm, creepy, likeable, polite & 27
    & \cite{andrews2012system}\cmt{[38]}
    \cite{ashktorab2019resilient}\cmt{[88]}
    \cite{ceha2022expressive}\cmt{[77]}
    \cite{ceha2021can}\cmt{[57]}
    \cite{chan2021kinvoices}\cmt{[74]}
    \cite{cox2022does}\cmt{[27]}
    \cite{cuadra2021my}\cmt{[67]}
    \cite{haas2022keep}\cmt{[78]}
    \cite{habler2019effects}\cmt{[63]}
    \cite{healey2013relating}\cmt{[39]}
    \cite{hu2022polite}\cmt{[76]}
    \cite{huiyang2022improving}\cmt{[17]}
    \cite{jeong2019exploring}\cmt{[10]}
    \cite{jestin2022effects}\cmt{[81]}
    \cite{khooshabeh2011does}\cmt{[37]}
    \cite{kim2019comparing}\cmt{[89]}
    \cite{lee2019s}\cmt{[55]}
    \cite{lee2020hear}\cmt{[23]}
    \cite{linnemann2018can}\cmt{[15]}
    \cite{miehle2018exploring}\cmt{[51]}
    \cite{moilanen2022measuring}\cmt{[82]}
    \cite{ouchi2019should}\cmt{[59]}
    \cite{tolmeijer2021female}\cmt{[62]}
    \cite{volkel2021manipulating}\cmt{[68]}
    \cite{volkel2022user}\cmt{[75]}
    \cite{wester2015artificial}\cmt{[14]}
    \cite{zhu2022effects}\cmt{[26]} 
\\ \Xhline{1.2pt}
\end{tabular}%
}
\caption{Perception measures}
\label{tab:perceptions}
\end{table*}

@@

\textbf{Perception of interaction with agent} evaluates the overall interaction quality users had with the conversational agent. The three basic themes for this perceptions are usability, engagement and satisfaction. \textit{Usability} measures the utilitarian aspects of the interaction, whether it was accurate, easy to use, efficient or helpful. Some commonly used methods to measure usability include the response accuracy portion of the SASSI questionnaire \cite{hone2000towards}\cmt{sassi}, or the NASA Task Load Index (NASA-TLX) \cite{hart1988development}\cmt{nasa} to measure cognitive workload. Questions such as "the system is easy to use" or "it is easy to understand the agent" are also measures of usability. \textit{Engagement} on the other hand measures users' emotional reactions, whether they have enjoyed the conversation with CA, or annoyed or frustrated with the interaction. Some commonly used methods to measure engagement include  the annoyance portion of the SASSI questionnaire \cite{hone2000towards}\cmt{sassi}, and the Use Engagement Scale (UES) \cite{o2018practical}\cmt{ues}. Questions such as "I enjoyed using the system" or "I felt frustrated with the agent" are also measures of engagement. Lastly, \textit{Satisfaction} measures users' overall satisfaction interacting with the agent. Questions such as "the overall assessment of conversing with the CA was satisfactory" are used to measure this perception.

\textbf{Perception of agent's ability} evaluates the perceived capabilities of the agent. While there are some elements of system performance in this measure, the main aspect we are interested in is the perception differences for agents with identify system capabilities but with different architecture elements. Specifically, \textit{intelligence} measures the agent's expertise and knowledge, and it is commonly administered through the Godspeed questionnaire \cite{bartneck2009measurement}\cmt{godspeed} on the set of questions related to perceived intelligence. Also survey questions such as asking users about the agent's intelligence and domain knowledge are used to measure perceived intelligence. \textit{Competence} takes intelligence one step further by evaluating the agent's ability to put its intelligence into practice. To measure a CA's competence, there are usually surveys or qualitative feedback that the agent is capable and competent, and that the users have confidence in the agent's ability to get the job done. Lastly, \textit{trust} evaluates agent's truthfulness, credibility and benevolence. Some commonly used methods to evaluate trust include the Trust Propensity Scale \cite{mayer1999effect} and the Individualized Trust Scale (ITS) \cite{wheeless1977measurement}. Questions such as "is the agent honest" and "can I trust the agent with sensitive information" are also measures of trust.

\textbf{Perception of the social connection with agent} evaluates the emotional connections that users have with agents. The three basic themes of conversation tone, social presence and intimacy are included in this category. \textit{Conversation tone} measure the affective impression of the agent's tone, such as empathy, expressiveness, or passionate. Also, it measures whether the tone used by the agent is persuasive or appropriate for the conversation. \textit{Social presence} measures the sense of connectness and psychological distance users has with an agent. Some aspects in this measure include a sense of familiarity or similarity with the agent, whether users feel the agent behaves like them or have similar attitudes to them. The measure of \textit{Intimacy} extends social presence into the realm of the quality of relationships with an agent. Some commonly used measures include the set of social attraction questions from the interpersonal attraction questionnaire \cite{mccroskey1975development} and the quality of relationship inventory (QRI) \cite{pierce1997assessing}. These questionnaires include questions like "I think the agent could be a friend of mine", or "I feel we could establish a personal relationship with each other".

\textbf{Perception of agent's humanness} evaluates human-likeness and the anthropomorphized personality traits that users assign to agents. \textit{Human-likeness} measures whether the agent presented itself as natural and human-like, or artificial and machine-like. The Godspeed questionnaire \cite{bartneck2009measurement}\cmt{godspeed} set of questions related to anthropomorphism and the Ascent of Man scale \cite{kteily2015ascent} are commonly used methods to assess human-likeness. Survey questions with semantic scales such as "human-like / machine-like" and "artificial / natural" are also used for this perception. \textit{Personality traits} captures the human characteristics that are attributes to the agent, such as warm, friendly, likeable, and polite. This is commonly captured in the qualitative feedback from users, commenting on whether the agent is extroverted or introverted, or the human characteristics they perceive the agent such as funny or witty. Some studies used the measure of the Big-5 personality traits \cite{gosling2003very} to map an agent's disposition on various personality dimensions.

@@

\begin{figure*}[]
  \centering
  \includegraphics[width=\textwidth]{fig-heatmap-impact.png}
  \caption{Heatmap of literature for perception of conversational agents for each type of conversational architecture element}
  \label{fig:heatmap-impact}
\end{figure*}

We extracted 265 connections between anthropomorphized perceptions and conversation architecture elements from our reviewed corpus. Out of these connections, 193 of them had an impacted relationship between the conversational architecture elements and anthropomorphized perceptions. Most of these relationships (n=183) were exploring the effect of a single conversation architecture element on perceptions of agents (e.g. \cite{miehle2018exploring}\cmt{[51]}\cite{westerman2019believe}\cmt{[9]}), while a few of them (n=10) used a composite of architecture elements in the study (e.g. \cite{seeger2021chatbots}\cmt{[35]}\cite{volkel2021manipulating}\cmt{[68]}). The 183 single element relationships are visualized into a heatmap (Figure \ref{fig:heatmap-impact}) to based on the number of impacts conversational architecture elements had on the anthropomorphized perceptions. These relationships are discussed in details in the following sections for each perception category.

%\subsubsection{Overall Characteristics}
